---
title: "treatment_effect"
author: "Alisaww"
date: "2022-07-28"
output: html_document
---
```{r}
library(dplyr)
library(ggplot2)
library(hexbin)
library(RColorBrewer)
library(extrafont)
library(ggfortify)
library(ggsci)
#library(forecast)
#library(zoo)
library(extrafont)
```


Firstly, load the data and see the effect of the drug treatment on the total PANSS score over time.
```{r}
getwd()
setwd("/Users/quyunjie/Desktop/STATS202/FinalProjectData")
A=read.csv("Study_A.csv")
B=read.csv("Study_B.csv")
C=read.csv("Study_C.csv")
D=read.csv("Study_D.csv")
E=read.csv("Study_E.csv")
summary(E)
```
```{r}
#remove LeadStatus column
A1 = A[ , -which(names(A) %in% c("LeadStatus"))]
B1 = B[ , -which(names(B) %in% c("LeadStatus"))]
C1 = C[ , -which(names(C) %in% c("LeadStatus"))]
D1 = D[ , -which(names(D) %in% c("LeadStatus"))]
combined=rbind(A1,B1,C1,D1,E)
```

First plot `PANSS_Total` vs `VisitDay`, for the `Treatment` and `Control` groups.
```{r}
treatment=combined[which(combined$TxGroup == "Treatment"),]
control=combined[which(combined$TxGroup == "Control"),]

p=ggplot(NULL,aes())+
  geom_point(data=treatment,aes(x=VisitDay,y=PANSS_Total,col="Treatment"),size=0.7)+
  geom_point(data=control,aes(x=VisitDay,y=PANSS_Total,col="Control",),shape=23,size=0.7)+
  #scale_color_manual(values=rev(pal_aaas("default")(11)))+
  #geom_line(data=treatment,aes(x=VisitDay,y=PANSS_Total,col="Treatment"))+
  #geom_line(data=control,aes(x=VisitDay,y=PANSS_Total,col="Control"))+
  theme_light()+
  theme(axis.title = element_text(color='black', vjust=0.1),
       axis.ticks.length = unit(-0.4,"lines"),
       axis.ticks = element_line(color='black'),
       legend.position ="top",
       legend.title=element_blank())
ggsave("pansstotal_vs_visitday.png",width=6,height=4,units="in",device="png",dpi="retina")

```

A plot of the total PANSS score over time does not seem to suggest a treatment effect. But to quantify this we can do a linear regression on the total PANSS score vs time, where we regress PANSS_Total onto VisitDay, and a dummy variable for Control or Treatment interacting with VisitDay: $y=\beta_0 + \beta_1*VisitDay + \beta_2*VisitDay*Treatment$

```{r}
i=38

p = ggplot(NULL,aes())+
  geom_point(data=treatment,aes(x=VisitDay,y=treatment[,i],col="Treatment"),size=0.5)+
  geom_point(data=control,aes(x=VisitDay,y=control[,i],col="Control"),size=0.5)+
  #scale_color_manual(values=rev(pal_aaas("default")(2)))+
  theme_minimal()+
  theme(legend.position = "top",legend.title=element_blank(),plot.title=element_text(hjust=0.5,size=10,family="Lato"),plot.subtitle=element_text(hjust=0.5,size=8,family="Lato"),text=element_text(size=10,family="Lato"))
print(p)
```
Again do not see an appreciable difference between control and treatment. 

```{r}
combined$totalP=rowSums(combined[,c("P1","P2","P3","P4","P5","P6","P7")])
combined$totalN=rowSums(combined[,c("N1","N2","N3","N4","N5","N6","N7")])
combined$totalG=rowSums(combined[,c("G1","G2","G3","G4","G5","G6","G7","G8","G9","G10","G11","G12","G13","G14","G15","G16")])
treatment=combined[which(combined$TxGroup == "Treatment"),]
control=combined[which(combined$TxGroup == "Control"),]
p = ggplot(NULL,aes())+
  geom_point(data=treatment,aes(x=VisitDay,y=totalG,col="Treatment"),size=0.5)+
  geom_point(data=control,aes(x=VisitDay,y=totalG,col="Control"),size=0.5)+
  #scale_color_manual(values=rev(pal_aaas("default")(2)))+
  theme_minimal()+
  theme(legend.position = "top",
        legend.title=element_blank(),
        plot.title=element_text(hjust=0.5,size=10,family="Lato"),
        plot.subtitle=element_text(hjust=0.5,size=8,family="Lato"),text=element_text(size=10,family="Lato"))
ggsave("panss_total_g.png",width=6,height=4,units="in",device="png",dpi="retina")
print(p)
#p + p+ facet_grid(cols = vars(3))
```

```{r}
combined$TxGroup = factor(combined$TxGroup)
lm.fit=lm(PANSS_Total~VisitDay+VisitDay:TxGroup, data=combined)
contrasts(combined$TxGroup) # display encoding for treatment 
```
The dummy encoding is 1 for treatment, 0 for Control.

```{r examine linear regression}
summary(lm.fit)
```
```{r}
lm.fit.p=lm(totalP~VisitDay+VisitDay:TxGroup, data=combined)
lm.fit.n=lm(totalN~VisitDay+VisitDay:TxGroup, data=combined)
lm.fit.g=lm(totalG~VisitDay+VisitDay:TxGroup, data=combined)
```

```{r}
summary(lm.fit.p)
summary(lm.fit.n)
summary(lm.fit.g)
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

